Investigating Determinism of Gaming Engines for Autonomous Vehicle Verification

### 2020-02-11 paper/1_Intro.tex
update to scheduling in background

% Verification and Determinism 
For game engines to be useful as a verification tool then they must be deterministic. 
%
% What is a deterministic system
Deterministic, in this context, refers to the property of causality given a temporal development of events such that any state is completely determined by prior states.
%
Herein the term \textit{simulation variance} is used to refer to how simulation runs vary when repeated many times and specifically the variance of agent paths within a given scenario. We will adopt the terminology defined in~\cite{Ulbrich2015}, where \textit{scene} refers to all static objects including the road network, street furniture, environment conditions and a snapshot of any dynamic elements. Dynamic elements are the elements in a scene whose actions or behaviour may change over time, these are considered actors and may include other road vehicles, cyclists, pedestrians and traffic signals. The \textit{scenario} is then defined as a temporal development between several scenes which may be specified by specific goals and values. A \textit{situation} is defined as the subjective conditions and determinants for behaviour at a particular point in time.
% 
% Why is determinism important for verification
Determinism is important for verification. Any software defects that are detected but not repeatable at a later time or by another party does not allow for suitable resolution. Furthermore, defects not found on the verification testing system may arise on a deployed system with a different hardware architecture resulting in untested and potentially unsafe vehicle behaviour. 
%
Also, if the simulation is non-deterministic, variance in actor positions may lead to subsequent assessment errors, e.g. assertion checking.

@@

This paper is structured as follows. In Section~\ref{s:background} a background on gaming engines and sources of potential non-determinism are explored. Section~\ref{s:methodology} presents a method to design and execute an experiment to assess suitability of a system for CAV verification. A case study to illustrate the level of simulation variance for a number of scenarios is explored in Section~\ref{s:case-study}. We conclude in Section~\ref{s:conclusion} and give an outlook on future work.

>>>>

% Verification and Determinism 
For game engines to be useful as a verification tool then they must be deterministic. 
%
% What is a deterministic system
Deterministic, in this context, refers to the property of causality given a temporal development of events such that any state is completely determined by prior states.
%
Herein the term \textit{simulation variance} is used to refer to how simulation outputs vary when repeated many times and specifically the variance of actor paths within a given scenario. We will adopt the terminology defined in~\cite{Ulbrich2015}, where \textit{scene} refers to all static objects including the road network, street furniture, environment conditions and a snapshot of any dynamic elements. Dynamic elements are the elements in a scene whose actions or behaviour may change over time, these are considered actors and may include other road vehicles, cyclists, pedestrians and traffic signals. The \textit{scenario} is then defined as a temporal development between several scenes which may be specified by specific goals and values. A \textit{situation} is defined as the subjective conditions and determinants for behaviour at a particular point in time.

% Why is determinism important for verification
Determinism is important for verification. Any software defects that are detected but not repeatable at a later time or by another party does not allow for suitable resolution. Furthermore, defects not found on the verification testing system may arise on a deployed system with a different hardware architecture resulting in untested and potentially unsafe behaviour. %When considering the development of autonomous vehicles, untested code could lead to dangerous behaviours not being identified and fatal road accidents may occur.
%
Also, if the simulation is non-deterministic, variance in actor positions may lead to subsequent assessment errors, e.g. assertion checking.

@@

This paper is structured as follows. In Section~\ref{s:background} a background on gaming engines and sources of potential non-determinism are explored. Section~\ref{s:methodology} presents a method to design and execute an experiment to assess suitability of a system for CAV verification. A case study to illustrate how the simulation variance vary for a number of scenarios is explored in Section~\ref{s:case-study}. We conclude in Section~\ref{s:conclusion} and give an outlook on future work.
